Grow with the flow: faster growth is associated with more variable precipitation in a perennial herb

### 2015-04-30 ms/ms.Rnw
Finished rough draft

<<Front Matter>>=

@@

\title{(working title) Local Physiological Adaptation in \textit{Mimulus cardinalis}}
\author{Christopher D. Muir and Amy L. Angert}
%\date{}							% Activate to display a given date or no date

@@

\begin{document}
\maketitle
\listoffigures
\listoftables

@@

Local adaptation is one of the most ubiquitous observations in nature: organisms perform well in their natal environment, but poorly outside it. Local adaptation within species most often involves genetic differences in ecologically relevant traits caused by selection to different environments, leading to predictable trait-environment correlations that persist in a common garden. Although local adaptation is well-supported by data from reciprocal transplant experiments, trait-environment associations, and ecological genetics/genomics, predicting which traits respond to selection exceeds our understanding. In particular, the role of physiological traits is poorly understood, even though physiology explains adaptation to the abiotic environment, which we need to understand as a baseline for anticipating how organisms will respond to climate change. A related problem is identifying the exact selective agents, such as abiotic factors like soil water availability or biotic interactions [etc.]... Here, we simultaneously investigate which traits and causal selective agents explain range-wide patterns of local physiological adaptation to climate in \textit{Mimulus}, a model genus for local adaptation studies.

@@

% Selective agents
We also know little about the selective agents, such as climatic factors or biotic interactions, that are most important in driving adaptive divergence. In certain situations, such as pollinator shifts and edaphic adaptation, it is straightforward to identify putative selective agents and test predicted mechanisms. However, it is more challenging for cases in which traits vary with less specific factors like mean annual precipitation or latitude. For example, body size clines (larger size at higher latitude) have evolved repeatedly on three contients in \textit{Drosophila subobscura}, strongly indicating selection. But which one or more of the many climatic factors that covary with latitude actually controls selection on body size? Selection might be even more complex for species that live over multiple seasons, meaning that environmental variation rather than the average environment may be a selective agent.

% Motivate using Mimulus
Linking physiological traits to potentially complex patterns of local adaptation requires integrating multiple lines of evidence from comparative, experimental, genomic studies under both lab and field conditions. Many classic and contemporary studies of local adaptation have been conducted using species from genus \textit{Mimulus} because of its natural history, easy propogation, and genetic/genomic resources. ... Conspicuous absence of phys mechanism

% Final paragraph
1. `intrinsic' physiological variation
2. variation in plasticity
3. what ecological factors explain divergence (i.e. putative selective agents underlying local adaptation)

Conclude that observations are best explained by stabilizing selection on the fundamental niche, but spatially varying divergent selection on coordinated life history and physiological traits.

@@

We used 16 populations from throughout the range of \textit{M. cardinalis} (Table 1). Seeds were collected in the field \cdm{Amy, can you explain seed collection methods?}.

<<focClim, echo = FALSE>>=
# Read in climate data for 16 focal populations to use in making Table 1
# [MAYBE SHOULD RENAMED "ExtendedTable1" or something?]
load(paste(pathDat, "/focClim.RData", sep = ""))

@@

\begin{table}[ht]
   \centering
   \topcaption{16 Focal populations}
   \begin{tabular}{@{} lllllll @{}}
      \toprule
  Name& Region  & Demo?& Pop gen?& Latitude  & Longtiude  & Elevation (mas) \\
      \midrule
	HAU & South Margin		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "HAU"], 3)}	& 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "HAU"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "HAU"]}   \\
	CTC	& South Margin		& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "CTC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "CTC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "CTC"]}   \\
	CUR	& South Margin		& yes	& yes & \Sexpr{round(focClim$Latitude[focClim$PopID == "CUR"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "CUR"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "CUR"]}   \\
	GRP & South Margin		& no	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "GRP"], 3)} &
    \Sexpr{round(focClim$Longitude[focClim$PopID == "GRP"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "GRP"]}   \\
	WWC &	Transverse		  & yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "WWC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "WWC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "WWC"]}   \\
	MIL	& Transverse	  	& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "MIL"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "MIL"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "MIL"]}   \\
	WFM	& Transverse	  	& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "WFM"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "WFM"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "WFM"]}   \\
	NMT	& South Sierras		& yes	& yes?& \Sexpr{round(focClim$Latitude[focClim$PopID == "NMT"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "NMT"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "NMT"]}   \\
	PRD	& South Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "PRD"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "PRD"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "PRD"]}   \\
	RWD	& South Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "RWD"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "RWD"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "RWD"]}   \\
	WNA	& Central Sierras	& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "WNA"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "WNA"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "WNA"]}   \\
	RBW	& Central Sierras	& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "RBW"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "RBW"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "RBW"]}   \\
	MYU	& North Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "MYU"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "MYU"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "MYU"]}   \\
	LIJ	& North Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "LIJ"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "LIJ"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "LIJ"]}   \\
	DPC	& North Coast	  	& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "DPC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "DPC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "DPC"]}   \\
	RCC	& North Margin		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "RCC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "RCC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "RCC"]}   \\
	\bottomrule
   \end{tabular}
\end{table}

@@

<<ExperimentalDesign>>=
library(scales) # for color transparency
exptlDes <- read.csv(paste(pathDat, "/ExperimentalDesign.csv", sep = ""))

@@

# Polygon of Drought treatment
# Rescale watering frequency to vary from 0 to 40
exptlDes$WaterFreq_dry <- exptlDes$WaterFreq_dry * (40 / 11) - 40 / 11

# with(exptlDes, points(DayN, WaterFreq_dry, type = "l", lwd = 12, 
	# col = alpha("white", 0.5)))
with(exptlDes, points(DayN, WaterFreq_dry, type = "l", lwd = 3, col = sec2shade4))

@@

# Legend
pdf(paste(pathFig, "/ExperimentalDesignLegend.pdf", sep = ""), 1.5, 2)
par(mai = rep(0, 4))
plot(0, 0, type = "n", xlim = c(0, 1.5), ylim = c(0, 1), axes = F)

polygon(c(0, 0.5, 1, 1, 0.5, 0), c(0.8, 1, 1, 0.55, 0.55, 0.35), border = NA, 
				col = alpha(primShade0, 0.25))
polygon(c(0, 0.5, 1, 1, 0.5, 0), c(0.6, 0.45, 0.45, 0, 0, 0.15), border = NA, 
				col = alpha(sec1shade0, 0.25))
points(c(0, 0.5, 1), c(0.575, 0.775, 0.775), lwd = 2, type = "l", col = primShade3)
points(c(0, 0.5, 1), c(0.375, 0.225, 0.225), lwd = 2, type = "l", col = sec1shade3)
text(0.5, 0.9, labels = "Hot", col = primShade3)
text(0.5, 0.1, labels = "Cool", col = sec1shade3)
text(1.25, 1, labels = "Max", col = alpha(primShade0, 0.25))
text(1.25, 0.55, labels = "Min", col = alpha(primShade0, 0.25))
text(1.25, 0.775, labels = "Mean", col = primShade3)

@@

\begin{figure}[h!]
	\centerline{\includegraphics[width=1\textwidth]{Figures/Figure_ExptlDes.pdf}}
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[Experimental Design]{CAPTION}
	\label{fig:Fig_ExptlDes}
\end{figure}

@@

\begin{table}[ht]
   \centering
   \topcaption{Traits}
   \begin{tabular}{@{} ll @{}}
      \toprule
  Trait & Units \\
      \midrule
  Day of germination  & day \\
  Leaf expansion rate  &  mm day$^{-1}$  \\
  Shoot elongation rate  &  mm or cm day$^{-1}$  \\
  Harvest dry mass &  g  \\
  Photosynthetic rate &  $\mu$mol CO$_2$ m$^{-2}$ s$^{-1}$\\
  Mortality &  Probability?  \\
	    \bottomrule
   \end{tabular}
   \label{table:Table_traits}
\end{table}

@@

<<Germination rate>>=

@@

<<Leaf expansion rate>>=

@@

<<Shoot elongation rate>>=

@@

<<Photosynthesis>>=

@@

<<Mortality>>=

@@

\subsection*{Principal components of germination, growth, and photosynthesis}
[maybe this goes in Results?]
We summarized population-level coefficients, after factoring out Treatment and other effects, using principal components. 
<<Principal Components>>=

@@

<<thinnedOccClimate>>=
# R data.frame of 178 M. cardinalis occurrences, thinned to correct for sampling bias, and associated climate data
load(paste(pathDat, "/thinnedOccClimate.RData", sep = ""))
@

@@

<<VSURF>>=

@@

For these analyses, we compiled a representative set of \Sexpr{nrow(thinnedOccClimate)} recent (since 2000) known \textit{M. cardinalis} occurences. These occurences were thinned by 50\% to correct for uneven sampling from a comprehensive set of herbarium records and an exhaustive field survey in 2010-11 \citep{Angert_ENM}. For each occurence, we used a 90m digital elevation model from HydroSHEDS \citep{Lehner_etal_2006} to extract elevation. Monthly interpolated climate layers were calculated using ClimateWNA \citep{Wang_etal_2012}, which accurately downscales climate data specifically for the rugged topography of western North America. For each occurence, we calculated bioclimatic variables using the biovars function in the R package \pkg{dismo} \citep{Hijmans_etal_2014}. In total, we included \Sexpr{length(climVars) / 2} cliamte variables, 9 from ClimateWNA and 15 bioclimatic variables (Table~\ref{table:TableS_ClimVars}). For each variable, we calculated both a 30-year normal by averging annual values between 1981 and 2010 and 30-year coefficient of variation, a standardized metric of interannual climatic variation. Temperatures were converted to Kelvin to be on a ratio scale appropriate for calculating the coefficient of variation. 
%--------------------------------------------------
% Results
%--------------------------------------------------

@@

4. Selective agent - unpredictability. Explain
Reiterate result
other supporting evidence: 
- ENM models are better for habitat suitability for S pops when using data from <5 years; better for Northern populations using 30 yr (M. Bayly \& A. Angert, unpub. data)
- greater variance in recruit size in S pops
-

There is a lot of interest in understanding how organisms will respond to changes in climatic variation, not just changes in the average climate. Our data indeed suggest that variation may be more important than the mean.

But what we really need to understand is why certain types of traits are repeatedly targets of natural selection, whereas other traits are constrained.

5. Future work and broad conclusions. Supports general conclusions that traits related to timing and growth are evolutionarily labile whereas those related to the fundamental niche are constrained (cites like \cite{Emery_etal_2012, Emery_Ackerly_2014}). Next, we are testing whether phenotypic variation and constraint can be predicted by the shape of tradeoffs, as predicted by evolutionary theory,

% stuff about environmental grain
How organisms adapt depends on the grain of the environment, ..., and gene flow between populations

Possibilities:
1. no trait variation: no environmental variation, high gene flow, generalist (convex)
2. variation in mean:
3. variation in plasticity: convex?,

our data are most consistent with:
- concave tradeoffs, fine-grained (coarse+concave would predict alt morphs), high gene flow

@@

\section*{Acknowledgements}
Erin Warkman and Lisa Lin helped collect data.

@@

%--------------------------------------------------
% Supplementary Information
%--------------------------------------------------

>>>>

<<Front Matter, echo=FALSE, results='hide'>>=

@@

\title{Grow with the flow: faster growth is associated with more variable precipitation in a perennial herb}
\author{Christopher D. Muir$^{1,*}$ and Amy L. Angert$^1$}
\date{}

@@

\begin{document}
\maketitle

$^1$ Biodiversity Research Centre, University of British Columbia, Vancouver, BC, Canada

$^*$corresponding author: Chris Muir, cdmuir@biodiversity.ubc.ca

Original Article

Key words: local adaptation, photosynthesis, growth rate, climatic variance, \textit{Mimulus}

Word counts: TBA

3 figures; 3 tables; ? supporting files

Data will be archived on Dryad upon acceptance.

\clearpage

\listoffigures
\listoftables

@@

Local adaptation is one of the most ubiquitous observations in nature: organisms perform well in their natal environment, but poorly outside it. Local adaptation within species most often involves genetic differences in ecologically relevant traits caused by selection to different environments, leading to predictable trait-environment correlations that persist in a common environment. Although local adaptation is well-supported by data from reciprocal transplant experiments, trait-environment associations, and ecological genetics/genomics, predicting which traits respond to selection is beyond our current understanding. In particular, the role of physiological traits is poorly understood, even though physiology explains adaptation to the abiotic environment, which we need to understand as a baseline for anticipating how organisms will respond to climate change. A related problem is identifying which features of the environment, abiotic factors like soil water availability or biotic interactions, cause spatially varying selective pressures. Here, we  investigate which traits and causal selective agents explain range-wide patterns of local physiological adaptation to climate in \textit{Mimulus}, a model genus for local adaptation studies.

@@

% Selective agents
We also know little about the selective agents, such as climatic factors or biotic interactions, that are most important in driving adaptive divergence. In certain situations, such as pollinator shifts and edaphic adaptation, it is straightforward to identify putative selective agents and test predicted mechanisms. However, it is more challenging for cases in which traits vary with less specific factors like mean annual precipitation or latitude. For example, body size clines (larger size at higher latitude) have evolved repeatedly on three contients in \textit{Drosophila subobscura}, strongly indicating selection \citep{Huey_etal_2000}. But which one or more of the many climatic factors that covary with latitude actually controls selection on body size? Selection might be even more complex for perennial species that live over multiple seasons, meaning that environmental variation rather than the average environment may be a selective agent.

% Motivate using Mimulus
Linking physiological traits to potentially complex patterns of local adaptation requires integrating multiple lines of evidence from comparative, experimental, genomic studies under both lab and field conditions. Many classic and contemporary studies of local adaptation have been conducted using species from genus \textit{Mimulus} because of its natural history, easy propogation, and genetic/genomic resources [willis, lowry?]. Yet, there is a conspicuous deficieny of links between local adaptation and physiological mechanisms \citep{Angert_2006, Angert_etal_2008}.

Two common signatures of local adaptation are nonrandom genetic differences in ecologically important traits and variation in plasticity (genotype by environment interaction). In this study, we found strong genetically-based differences in photosynthesis and growth, but no evidence for variation in plasticity in response to temperature and drought treatments. Interannual variation in precipitaiton is associated intrinsic variation in these traits, suggesting that climatic variance rather than mean may be an important driver of local adaptation in \textit{M. cardinalis}.

@@

We used 16 populations from throughout the range of \textit{M. cardinalis} (Table~\ref{table:Table_FocalPops}). Seeds were collected in the field \cdm{Amy, can you explain seed collection methods?}.

<<focClim, echo = FALSE, results='hide', warning=FALSE>>=
# Read in climate data for 16 focal populations to use in making Table 1
# [MAYBE SHOULD RENAMED "ExtendedTable1" or something?]
load(paste(pathDat, "/focClim.RData", sep = ""))

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table of focal population
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[ht]
   \centering
   \topcaption[Focal Populations]{Geographic region, latitude, longitude, and elevation (mas = meters above seal level) of 16 focal populations used in this study.}
   \begin{tabular}{@{} lllllll @{}}
      \toprule
  Name& Region  & Demo?& Pop gen?& Latitude  & Longtiude  & Elevation (mas) \\
      \midrule
	HAU & South Margin		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "HAU"], 3)}	& 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "HAU"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "HAU"]}   \\
	CTC	& South Margin		& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "CTC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "CTC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "CTC"]}   \\
	CUR	& South Margin		& yes	& yes & \Sexpr{round(focClim$Latitude[focClim$PopID == "CUR"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "CUR"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "CUR"]}   \\
	GRP & South Margin		& no	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "GRP"], 3)} &
    \Sexpr{round(focClim$Longitude[focClim$PopID == "GRP"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "GRP"]}   \\
	WWC &	Transverse		  & yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "WWC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "WWC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "WWC"]}   \\
	MIL	& Transverse	  	& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "MIL"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "MIL"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "MIL"]}   \\
	WFM	& Transverse	  	& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "WFM"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "WFM"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "WFM"]}   \\
	NMT	& South Sierras		& yes	& yes?& \Sexpr{round(focClim$Latitude[focClim$PopID == "NMT"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "NMT"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "NMT"]}   \\
	PRD	& South Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "PRD"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "PRD"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "PRD"]}   \\
	RWD	& South Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "RWD"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "RWD"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "RWD"]}   \\
	WNA	& Central Sierras	& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "WNA"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "WNA"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "WNA"]}   \\
	RBW	& Central Sierras	& yes	& no	& \Sexpr{round(focClim$Latitude[focClim$PopID == "RBW"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "RBW"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "RBW"]}   \\
	MYU	& North Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "MYU"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "MYU"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "MYU"]}   \\
	LIJ	& North Sierras		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "LIJ"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "LIJ"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "LIJ"]}   \\
	DPC	& North Coast	  	& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "DPC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "DPC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "DPC"]}   \\
	RCC	& North Margin		& yes	& yes	& \Sexpr{round(focClim$Latitude[focClim$PopID == "RCC"], 3)} & 
    \Sexpr{round(focClim$Longitude[focClim$PopID == "RCC"], 3)}	& \Sexpr{focClim$Elevation[focClim$PopID == "RCC"]}   \\
	\bottomrule
	\end{tabular}
	\label{table:Table_FocalPops}
\end{table}

@@

<<ExperimentalDesign, echo=FALSE, results='hide', warning=FALSE>>=
library(scales) # for color transparency
exptlDes <- read.csv(paste(pathDat, "/ExperimentalDesign.csv", sep = ""))

@@

# Line of Drought treatment
# Rescale watering frequency to vary from 0 to 40
exptlDes$WaterFreq_dry <- exptlDes$WaterFreq_dry * (40 / 11) - 40 / 11

# with(exptlDes, points(DayN, WaterFreq_dry, type = "l", lwd = 12, 
	# col = alpha("white", 0.5)))
with(exptlDes, points(DayN, WaterFreq_dry, type = "l", lwd = 3, col = "black"))

@@

# Legend
pdf(paste(pathFig, "/ExperimentalDesignLegend.pdf", sep = ""), 2, 2)
par(mai = rep(0, 4))
plot(0, 0, type = "n", xlim = c(-0.5, 1.5), ylim = c(0, 1), axes = F)

# Temperature
polygon(c(0, 0.5, 1, 1, 0.5, 0), c(0.8, 1, 1, 0.55, 0.55, 0.35), border = NA, 
				col = alpha(primShade0, 0.25))
polygon(c(0, 0.5, 1, 1, 0.5, 0), c(0.6, 0.45, 0.45, 0, 0, 0.15), border = NA, 
				col = alpha(sec1shade0, 0.25))
points(c(0, 0.5, 1), c(0.575, 0.775, 0.775), lwd = 2, type = "l", col = primShade3)
points(c(0, 0.5, 1), c(0.375, 0.225, 0.225), lwd = 2, type = "l", col = sec1shade3)

# Drought
lines(c(0, 1), c(1, 0), lwd = 2)
text(-0.25, 0.95, labels = "Water\nFreq")
text(0.5, 0.9, labels = "Hot", col = primShade3)
text(0.5, 0.1, labels = "Cool", col = sec1shade3)
text(1.25, 1, labels = "Max", col = alpha(primShade0, 0.25))
text(1.25, 0.55, labels = "Min", col = alpha(primShade0, 0.25))
text(1.25, 0.775, labels = "Mean", col = primShade3)

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure summarizing experimental design
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
	\centerline{\includegraphics[width=1\textwidth]{Figures/Figure_ExptlDes.pdf}}
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[Experimental Design]{Overview of experimental treatments and timing of key trait measurements. All plants germinated within 21 days of sowing. At that time, we began temperature treatments (left axis), simulating a typical June-August weather pattern at Hot (red) and Cold (blue) sites. The bold lines track the average daily temperatures. Within each day, there was a maximum daytime temperture (top of translucent polygons) and minimum nighttime temperature (bottom of translucent polygons). The drought treatment commenced later by ramping down the frequency of bottom-watering episodes (black line; right axis). Grey boxes on the bottom of the plot outline the period of key measurements described in the Methods.}
	\label{fig:Fig_ExptlDes}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table of key traits
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[ht]
   \centering
   \topcaption{Key traits measured in this study.}
   \begin{tabular}{@{} ll @{}}
      \toprule
  Trait & Units \\
      \midrule
  Day of germination  & day \\
  Leaf expansion rate  &  mm day$^{-1}$  \\
  Shoot elongation rate  &  mm or cm day$^{-1}$  \\
  Harvest dry mass &  g  \\
  Photosynthetic rate &  $\mu$mol CO$_2$ m$^{-2}$ s$^{-1}$\\
  Mortality &  Probability?  \\
	    \bottomrule
   \end{tabular}
   \label{table:Table_traits}
\end{table}

@@

<<Germination rate, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=

@@

<<Leaf expansion rate, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=

@@

<<Shoot elongation rate, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=

@@

<<Photosynthesis, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=

@@

<<Mortality, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=

@@

\subsection*{Principal components of germination, growth, and photosynthesis}
We summarized population-level coefficients, after factoring out Treatment and other effects, using principal components. The first principal component of these traits (Trait PC1) loaded positively with germination, growth, and photostynthetic rate, therefore we define this as a phenotypic axis dilineating fast and slow growing populations.

<<Principal Components, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=

@@

<<thinnedOccClimate, echo=FALSE, results='hide', warning=FALSE>>=
# R data.frame of 178 M. cardinalis occurrences, thinned to correct for sampling bias, and associated climate data
load(paste(pathDat, "/thinnedOccClimate.RData", sep = ""))
@

@@

<<VSURF, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=

@@

For these analyses, we compiled a representative set of \Sexpr{nrow(thinnedOccClimate)} recent (since 2000) known \textit{M. cardinalis} occurences. These occurences were thinned by 50\% to correct for uneven sampling from a comprehensive set of herbarium records and an exhaustive field survey in 2010-11 \citep{Angert_ENM}. For each occurence, we used a 90m digital elevation model from HydroSHEDS \citep{Lehner_etal_2006} to extract elevation. Monthly interpolated climate layers were calculated using ClimateWNA \citep{Wang_etal_2012}, which accurately downscales climate data specifically for the rugged topography of western North America. For each occurence, we calculated bioclimatic variables using the biovars function in the R package \pkg{dismo} \citep{Hijmans_etal_2014}. In total, we included \Sexpr{length(climVars) / 2} cliamte variables, 9 from ClimateWNA and 15 bioclimatic variables (Table~\ref{table:TableS_ClimVars}). For each variable, we calculated both a 30-year normal by averging annual values between 1981 and 2010 and 30-year coefficient of variation, a standardized metric of interannual climatic variation. Temperatures were converted to Kelvin to be on a ratio scale appropriate for calculating the coefficient of variation.

%--------------------------------------------------
% Results
%--------------------------------------------------

@@

Based on the available data, interannual variation in annual or winter precipitation (these are closely correlated in Mediterranean climates) may be the selective agent driving variation in growth and photosynthesis. Variation in precipitation was best predicted latitude of recent \textit{M. cardinalis} occurences and trait variation along the fast-slow growth continuum (Fig.~\ref{fig:Fig_ClimVarImp}). A life history tradeoff between allocation to growth in the current year at the expense of future years could explain this pattern. In southern populations with more frequent droughts capable of killing rhizomes, a more annualized strategy could be favored. Conversely, in more predictable northern environments, lifetime fitness may be optimized when a significant fraction of assimilate is allocated below ground for future years. Although this hypothesis remains to be directly tested, a few indepdent lines of evidence are consistent with it. Preliminary surveys suggest that northern populations not only grow slower, but also produce greater numbers of rhizomes (C.D. Muir, unpub. data), suggesting an allocation tradeoff. Ecological niche models also show that occurence of southern populations is best predicted by recent climate (< 5 years), whereas northern occurences are best predicted by climate over the previous 30 years (M. Bayly \& A. Angert, unpub. data). Finally, demographic surveys of natural populations show greater variation in the size recruits in southern populations, suggesting higher maximum growth rates under natural conditions (M. Bayly \& A. Angert, unpub. data). There is a lot of interest in understanding how organisms will respond to changes in climatic variation, not just changes in the average climate. Our data indeed suggest that variation may be more important than the mean.

[Future work and broad conclusions]. Supports general conclusions that traits related to timing and growth are evolutionarily labile whereas those related to the fundamental niche are constrained (cites like \cite{Emery_etal_2012, Emery_Ackerly_2014}). Next, we are testing whether phenotypic variation and constraint can be predicted by the shape of tradeoffs, as predicted by evolutionary theory. \cdm{I will work more on this once I am more comfortable with the rest of the discussion}.

@@

\section*{Acknowledgements}
Erin Warkman and Lisa Lin helped collect data. CDM was supported by a Biodiversity Postdoctoral Fellowship funded by the NSERC CREATE program. \cdm{Please fill in other relevant funding info}

@@

\clearpage

%--------------------------------------------------
% Supporting Information
%--------------------------------------------------

\section*{Supporting Information}

% Modify and restart table/figure numbering for appendixes
\renewcommand\thefigure{S\arabic{figure}}    
\renewcommand\thetable{S\arabic{table}}    
\renewcommand\theequation{S\arabic{equation}}    
\setcounter{table}{0}    
\setcounter{equation}{0}
\setcounter{figure}{0}
